---
title: "Marketing Campaign"
author: "Sri Harika Cherukuri"
date: "27/11/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(ggplot2)
```


```{r}
data = read_delim("bank-additional-full.csv", delim=";")
head(data,10)
```


```{r}
summary(data)
```

```{r}
sum(is.na(data))
any(is.null(data))
any(is.na(data))
```
```{r}

ggplot(data %>%
         count(y),
       aes(y,n,fill=y))+
  geom_bar(stat="identity")+
  labs(title = "Proportion of Customers who have opted for Term Deposits ", x = "If the Customer opted for term deposit",y="Number of Customers")+
  theme_minimal()+
  scale_x_discrete(labels = c("No","Yes"))+
  scale_y_continuous(labels = scales::number_format(big.mark = ','))+
  geom_text(aes(y = ((n)/sum(n)), label = scales::percent((n)/sum(n))),vjust = -0.25) +
  theme(legend.position = "none") 

```


```{r}

ggplot(data %>%
         count(age,y)%>%
         mutate(pct=n/sum(n)),
       aes(age,n,fill=y))+
  
  geom_bar(stat="identity")+
  labs(title = "Proportion of Customers who have opted for Term Deposits \nincreases with increase in Age", x = "Age of the Customer (in years)",y="Number of Customers")+
  guides(fill=guide_legend(title="Has the Customer \nopted for term deposit ?"))+
  theme_minimal()

```


```{r}
data %>% count(age)

ggplot(data)+
  geom_histogram(mapping = aes(x = age,fill=y),binwidth = 10,position="dodge")+
  labs(title = "Proportion of Customers who have opted for Term Deposits \nincreases with increase in Age", x = "Age of the Customer (in years)",y="Number of Customers")+
  guides(fill=guide_legend(title="Has the Customer \nopted for term deposit ?"))
```

```{r}
nrow(data%>%filter(age=="unknown"))/nrow(data)*100
```

```{r job}
data %>% count(job)

ggplot(data, 
       aes(x = forcats::fct_infreq(job), 
           fill = y)) + 
  geom_bar()+ 
  labs(title="Banks majorly reached out to customers who are \nAdmins, Blue-Collar and Technician ",
       x="Job Category",
       y="Number of Customers")+
  theme_minimal()+
  scale_y_continuous(labels = scales::number_format(big.mark = ','))+
  theme(axis.text.x=element_text(angle=50, hjust=1))+
  guides(fill=guide_legend(title="Has the Customer \nopted for term deposit ?"))
```

```{r}
nrow(data%>%filter(job=="unknown"))/nrow(data)*100
```


Most people contacted during the campaigns are from the "admin" job category and they are the ones who are highest in number of people who agreed for a term deposit.\

we can replace job values with retired where age is greater than or equal to 60.\

Rest of the rows with unknown job values can be dropped off.


```{r}
#unknown job values before imputation
data %>% filter(job=="unknown")
data$job[data$job=="admin."] <- "admin"
data$job[data$job=="unknown" & data$age>=60] <- "retired"
#unknown job values after imputation
data %>% filter(job=="unknown")
data = data %>% filter(job!="unknown")
```

```{r marital}
data %>% count(marital)

ggplot(data, 
       aes(x = forcats::fct_infreq(marital), fill=y)) + 
  geom_bar()+ 
  stat_count(aes(label=..count..), 
             vjust=0,
             geom="text", 
             position="identity")+
  theme(axis.text.x=element_text(angle=50, hjust=1))+
  labs(title="Marital Status",
       x="Status",
       y="Number of People")
```

```{r}
nrow(data%>%filter(marital=="unknown"))/nrow(data)*100

```

Dropping off the rows with unknown Marital Status values:

```{r}
data = data %>% filter(marital!="unknown")
```


```{r education}
data %>% count(education)

ggplot(data, 
       aes(x = forcats::fct_infreq(education), fill=y)) + 
  geom_bar()+ 
  stat_count(aes(label=..count..), 
             vjust=0,
             geom="text", 
             position="identity")+
  theme(axis.text.x=element_text(angle=50, hjust=1))+
  labs(title="Distribution of people in various education levels",
       x="Level of Education",
       y="Number of People")
```

```{r}
nrow(data%>%filter(education=="unknown"))/nrow(data)*100
```

Around 4% unknown values for education.\
Hypothesis is that the job will be related to the level of education, so we can fill out the education level with the help of job positions.


```{r}
ggplot(data, 
       aes(x = job, 
           fill = education)) +
  geom_bar(position = "stack")+
   labs(title = "Majority of the working customers seem to hold \na Univeristy Degree", x = "Job Category",y="Number of Customers")+
  theme_minimal()+ 
  scale_y_continuous(labels = scales::number_format(big.mark = ','))+
  theme(axis.text.x=element_text(angle=50, hjust=1))
```


Most occurring level of education in various jobs :

admin -> university.degree\
blue-collar -> basic.9y\
housemaid -> basic.4y\
management -> university.degree\
services -> high.school\
technician -> professional.course\

We can insert some education level values according to jobs:

```{r}
#unknown education before imputation

data %>% filter(education=="unknown")
  
data$education[data$education=="unknown" & data$job=="admin"] <- "university.degree"

data$education[data$education=="unknown" & data$job=="blue-collar"] <- "basic.9y"

data$education[data$education=="unknown" & data$job=="housemaid"] <- "basic.4y"

data$education[data$education=="unknown" & data$job=="management"] <- "university.degree"

data$education[data$education=="unknown" & data$job=="services"] <- "high.school"

data$education[data$education=="unknown" & data$job=="technician"] <- "professional.course"


#unknown education after imputation
data %>% filter(education=="unknown")
```


1613-387 = 1226 values imputed, 387 will be dropped from the education column.\

Now we can drop off the remaining rows with unknown education values:

```{r}
data = data %>% filter(education!="unknown")
nrow(data%>%filter(education=="unknown"))/nrow(data)*100
```


```{r default}
data %>% count(default)

ggplot(data, 
       aes(x = forcats::fct_infreq(default))) + 
  geom_bar()+ 
  stat_count(aes(label=..count..), 
             vjust=0,
             geom="text", 
             position="identity")+
  theme(axis.text.x=element_text(angle=50, hjust=1))+
  labs(title="Have people defaulted ?",
       x="Defaulted - Yes and No",
       y="Number of People")
```

```{r}
nrow(data%>%filter(default=="unknown"))/nrow(data)*100
```


```{r}
# data$default[data$default=="unknown"] <- "no"
nrow(data%>%filter(default=="unknown"))/nrow(data)*100
```

```{r housing}
data %>% count(housing)

ggplot(data, 
       aes(x = forcats::fct_infreq(housing), fill=y)) + 
  geom_bar()+ 
  stat_count(aes(label=..count..), 
             vjust=0,
             geom="text", 
             position="identity")+
  theme(axis.text.x=element_text(angle=50, hjust=1))+
  labs(title="Has the person taken housing loan ?",
       x="Loan taken ?",
       y="Number of People")
```

```{r}
nrow(data%>%filter(housing=="unknown"))/nrow(data)*100
```


```{r}
data = data %>% filter(housing!="unknown")
nrow(data%>%filter(housing=="unknown"))/nrow(data)*100
```

```{r loan-personal}
data %>% count(loan)

ggplot(data, 
       aes(x = forcats::fct_infreq(loan), fill=y)) + 
  geom_bar()+ 
  stat_count(aes(label=..count..), 
             vjust=0,
             geom="text", 
             position="identity")+
  theme(axis.text.x=element_text(angle=50, hjust=1))+
  labs(title="Has the person taken personal loan ?",
       x="Loan taken ?",
       y="Number of People")
```

```{r}
nrow(data%>%filter(loan=="unknown"))/nrow(data)*100

```

Unknown values in the housing and personal loan columns correspond the same rows.

```{r contact}
data %>% count(contact)

ggplot(data, 
       aes(x = forcats::fct_infreq(contact), fill=y)) + 
  geom_bar()+ 
  stat_count(aes(label=..count..), 
             vjust=0,
             geom="text", 
             position="identity")+
  theme(axis.text.x=element_text( hjust=0.5))+
  labs(title="People Contacted over different mediums",
       x="Contact Medium ?",
       y="Number of People")


```

```{r month}
data %>% count(month)

ggplot(data, 
       aes(x = forcats::fct_infreq(month), fill=y)) + 
  geom_bar()+ 
  stat_count(aes(label=..count..), 
             vjust=0,
             geom="text", 
             position="identity")+
  theme(axis.text.x=element_text( hjust=0.5))+
  labs(title="People Contacted over different months",
       x="Month",
       y="Number of People")


```


```{r day-of-week}
data %>% count(day_of_week)

ggplot(data, 
       aes(x = forcats::fct_infreq(day_of_week), fill=y)) + 
  geom_bar()+ 
  stat_count(aes(label=..count..), 
             vjust=0,
             geom="text", 
             position="identity")+
  theme(axis.text.x=element_text( hjust=0.5))+
  labs(title="People Contacted over different days of the week",
       x="Week Day",
       y="Number of People")

```


```{r duration}
data %>% count(duration)

nrow(data%>%filter(duration==0))/nrow(data)*100
```

First "yes" response was around 37 seconds call duration.

```{r}
data %>% filter(y=="yes" & duration==37)
# temp = data %>% group_by(campaign,y ) %>% count()
# data = data %>% filter(duration>10)
# summary(data$duration)
```

Scatter plot - duration vs number of people:

```{r}
data %>% group_by(duration) %>% count() %>%
ggplot(aes(x=duration,y=n))+
  geom_point(size=0.2)+
  geom_jitter()
```

```{r}
ggplot(data, aes(x=duration, fill=y))+
  geom_histogram(mapping = aes(y = after_stat(log10(count))))+
  labs(x="Duration of Call (in seconds)", y="Log base 10 Transoformation \nof Number of Customers",title= "Proportion of Customers who have opted for Term Deposits \nincreases with duration of call")+
  xlim(0,2500)+
  guides(fill=guide_legend(title="Has the Customer \nopted for term deposit ?"))+
  theme_minimal()

```
  
```{r campaign}
data %>% count(campaign)
```

```{r}
ggplot(data, aes(x=campaign, fill=y))+
  geom_histogram()+
  xlim(0,20)
```

Checking proportion of people converted over number of campaigns.

```{r}
temp = data %>% group_by(campaign,y ) %>% count()
temp = temp %>% pivot_wider(names_from = y, values_from = n)
temp %>% mutate(p=yes/(yes+no))

```

```{r}
data %>% filter(campaign==56)
```

The above person was contacted 56 times during previous campaigns but
the call duration was 261 seconds and final response was NO.

```{r pdays}
data %>% count(pdays)

nrow(data%>%filter(pdays==999))/nrow(data)*100
```
Almost 96% of the customers haven't been contacted previously.

```{r}

temp1 = data %>% filter(pdays!=999)
ggplot(temp1, aes(x=pdays, fill=y))+
  geom_histogram()+
  labs(title="Number of people distributed across number of days passed since the\n last call excluding first time customers")

```

```{r}
temp2 = temp1 %>% group_by(pdays,y ) %>% count()
temp2 = temp2 %>% pivot_wider(names_from = y, values_from = n)
temp2 %>% mutate(p=yes/(yes+no))
```

```{r previous}
data %>% count(previous)

```

```{r}
ggplot(data, aes(x=previous, fill=y))+
  geom_histogram()
```

```{r poutcome}
data %>% count(poutcome)

ggplot(data, 
       aes(x = forcats::fct_infreq(poutcome), fill=y)) + 
  geom_bar()+ 
  stat_count(aes(label=..count..), 
             vjust=0,
             geom="text", 
             position="identity")+
  theme(axis.text.x=element_text( hjust=0.5))+
  labs(title="Result from previous campaign",
       x="Result",
       y="Number of People")

```

Number of 'nonexistent' entries in 'poutcome' = 34122 = number of people who have been contacted previously 0 times

```{r cons.price.idx}
data %>% count(cons.price.idx)
summary(data$cons.price.idx)
```

```{r}
ggplot(data, aes(x=cons.price.idx, y))+
  geom_boxplot()
```

High price index -> people setting up the term deposit
Low price index -> people not setting up the term deposit

```{r cons.conf.idx}
data %>% count(cons.conf.idx)
```

```{r}
ggplot(data, aes(x=cons.conf.idx, y))+
  geom_boxplot()
```

High confidence index -> people setting up the term deposit
Low conf index -> people not setting up the term deposit

```{r emp.var.rate}
data %>% count(emp.var.rate)
```

```{r}
ggplot(data, aes(x=emp.var.rate, fill=y))+
  geom_histogram(binwidth = 1)
```

```{r nr.eployed}
data %>% count(nr.employed)
```
Employment Variation is low and its a quarterly indicator.

```{r euribor3m}
data %>% count(euribor3m)
```

```{r}
ggplot(data, aes(x=euribor3m, y))+
  geom_boxplot()
```

High Interest -> didn't set up the term deposit
Low Interest -> set up the term deposit

Getting the filtered data frame:
```{r}
filtered = data

#write.csv(filtered,"bank_filtered.csv", row.names = FALSE)
filtered
```

Heat Map:
```{r}

data1 = select_if(data, is.numeric)

data$y[data$y=="yes"] <- 1
data$y[data$y=="no"] <- 0

data1$y = as.numeric(as.character(data$y))
```

```{r}
library(reshape2)
cor_matrix = round(cor(data1),2)
melted_matrix <- melt(cor_matrix)

ggplot(melted_matrix, aes(x=Var1, y=Var2, fill= value)) +
  geom_tile()+
geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
theme(
  axis.text.x=element_text(angle=50, hjust=1),
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  legend.direction = "vertical")

```

## Additional
```{r}
# Duration vs Campaign
data %>%
ggplot(aes(x=duration/60,y=campaign,color = y))+
  geom_jitter(size=0.5)+
  xlab("Duration of Calls in Minutes")+
  ylab("Number of Calls in this Campaign")+
  ggtitle("Duration vs Number of Calls")+
  theme_bw()
```

## Correlation check

### Categorical variables
```{r}

library(dplyr)

col = c("job","marital","education","default","housing","loan","contact","month","day_of_week","poutcome","y")

for (i in col){
  for (j in col){
    print(paste("P-value for ",i, "and",j,"is", round(chisq.test(get(i,data),get(j,data),simulate.p.value = TRUE)$p.value,5)))
  }
}
```
If we consider a threshold of 0.05 for p-value: loan can be removed.
```{r}
#install.packages("corrplot")
library(corrplot)
for (i in col){
  c <- chisq.test(get(i,data),data$y,simulate.p.value = TRUE)
  corrplot(c$residuals,is.cor=FALSE)
}
```
http://www.sthda.com/english/wiki/chi-square-test-of-independence-in-r
https://www.mathsisfun.com/data/chi-square-test.html
Positive residuals are in blue. Positive values in cells specify an attraction (positive association) between the corresponding row and column variables.\
Negative residuals are in red. This implies a repulsion (negative association) between the corresponding row and column variables.

### Continuous variables
```{r}
col = c("nr.employed","euribor3m","cons.conf.idx","cons.price.idx","emp.var.rate","pdays","campaign","duration","age")

con = data[,col]
cor_matrix = round(cor(con),2)
melted_matrix <- melt(cor_matrix)

ggplot(melted_matrix, aes(x=Var1, y=Var2, fill= value)) +
  geom_tile()+
geom_text(aes(Var2, Var1, label = value), color = "white", size = 4) +
theme(
  axis.text.x=element_text(angle=50, hjust=1),
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  legend.direction = "vertical")
```
euribor3m, emp.var.rate, nr.employed are highly correlated with each other. CPI is also highly correlated with these three >0.5.

### Continuous and Categorical variables
```{r}
col = c("nr.employed","euribor3m","cons.conf.idx","cons.price.idx","emp.var.rate","pdays","campaign","duration","age")

results <- purrr::map(data[,col],~aov(.x~data$y))

m=1
for (i in col){

  pval <- unlist(summary(results[[m]]))
  print(paste("P value with ",i,"is",pval["Pr(>F)1"][[1]]))
  m=m+1
  
}


```

Continuous and Categorical variables - using pearson. which is known as point- biserial since the response variable will be encoded

```{r}

cor_matrix = round(cor(data1[,c(col,'y')]),2)
melted_matrix <- melt(cor_matrix)

ggplot(melted_matrix, aes(x=Var1, y=Var2, fill= value)) +
  geom_tile()+
geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
theme(
  axis.text.x=element_text(angle=50, hjust=1),
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  legend.direction = "vertical")

```
Duration can be considered.
nr.emploed, cpi, emp.var.rate,euribor are dependent of eac other- So, nr.employed can be considered since it has relatively larger coefficient.
(let's also try nr.eployed with cpi)
p days can also be considered. Rest of the coeffiecients seem sma;ll

### Unbalanced Data
```{r}
#install.packages("RColorBrewer")
library(RColorBrewer)

color <- brewer.pal(length(count), "Set2") 

pi <- data1 %>% group_by(y) %>% count()
pie(pi$n,labels=paste(pi$y,"=",round(100*pi$n/sum(pi$n),2),"%"),col=color)
```

The algorithm receives significantly more examples from one class, prompting it to be biased towards that particular class. It does not learn what makes the other class “different” and fails to understand the underlying patterns that allow us to distinguish classes.

To treat this we better proceed with synthetic data generation- SMOTE https://www.analyticsvidhya.com/blog/2016/03/practical-guide-deal-imbalanced-classification-problems/


